---
title: "DS510-Final"
author: "Team13"
date: "2023-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE}
library(readxl)
library(dplyr)
library(ggplot2)
library(reshape2)
library(caret)
library(pROC)
```


```{r}


df <- read_excel("default of credit card clients.xls", skip = 1)
df <- df %>% rename(default_pay = `default payment next month`)
```

```{r}
bins <- c(20,30,40,50,60,70,80)
names <- c('21-30','31-40','41-50','51-60','61-70','71-80')
df$age_bin <- cut(x=df$AGE, breaks=bins, labels=names, right=TRUE)

age_cnt <- data.frame(table(df$age_bin))
age_0 <- data.frame(table(df$age_bin[df$default_pay == 0]))
age_1 <- data.frame(table(df$age_bin[df$default_pay == 1]))

ggplot() + 
  geom_bar(data = age_0, aes(x = Var1, y = Freq), stat = "identity", fill = "red", alpha = 0.5) +
  geom_bar(data = age_1, aes(x = Var1, y = Freq), stat = "identity", fill = "blue", alpha = 0.5) +
  labs(title = "Number of clients in each age group", x = "Age group", y = "Number of clients") +
  scale_x_discrete(labels = c("21-30", "31-40", "41-50", "51-60", "61-70", "71-80")) +
  theme_minimal()

```

```{r}
library(dplyr)
rename_df <- df %>%
   mutate(default_pay = recode(default_pay, "1" = "Yes", "0" ="No"))
ggplot(rename_df, aes(x=default_pay, y=AGE, fill=default_pay)) +
   geom_boxplot() +
   scale_fill_brewer(palette="Set1", direction=-1) +
   labs(y="Age",
        x="Default pay",
        title="Age associated with default pay") +
   theme_minimal()
```


```{r}
corrdata <- df %>% select(LIMIT_BAL, SEX, EDUCATION, MARRIAGE, AGE, PAY_0, PAY_2, PAY_3, PAY_4, PAY_5, PAY_6, default_pay)
corr <- cor(corrdata)
melted_corr <- melt(corr)

ggplot(melted_corr, aes(x=Var1, y=Var2, fill=value)) +
geom_tile() +
 scale_fill_gradient2(low="blue", mid="white", high="red", midpoint=0) +
   geom_text(aes(label=round(value,2)), color="black", size=3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45, vjust=1, 
                                    size=12, hjust=1),
        axis.text.y = element_text(size=12),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size=10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
```

From this correlated plot, we could see that PAY_5 to PAY_6 have a highest correlation to each other(0.82). We should also consider to remove features with a higher correlation to avoid overfitting. According to above information, I'm going to remove PAY_5 at this point.


```{r}
# drop PAY_5
df <- df[, !(names(df) %in% c('PAY_5'))]

# encode
df <- df %>% 
  mutate_at(vars(EDUCATION, MARRIAGE,default_pay), factor)

```

```{r}
set.seed(1)
train <- createDataPartition(df$default_pay, p=0.7, list=FALSE)

table(df$default_pay[train])
df_train <- df[as.integer(train),]
df_test <- df[-as.integer(train),]


fit <- glm(default_pay ~ ., data=df_train, family=binomial(link="logit"))

summary(fit)

```

```{r}
prob <- predict(fit, newdata=df_test, type="response")

pred <- ifelse(prob > 0.5, 1, 0)

hist(prob)

```

```{r}
#  calculate the confusion matrix and statistics
cm <- confusionMatrix(factor(pred), df_test$default_pay, positive="1")
cm
```

```{r}
test_prob <- predict(fit, newdata = df_test, type = "response")
test_roc <- roc(df_test$default_pay , test_prob, plot = TRUE, print.auc = TRUE)
```


```{r}

ggplot(df, aes(x = default_pay)) +
  geom_bar(fill = "steelblue") +
  scale_fill_brewer(palette="Set1", direction=-1) +
  labs(title = "Default payment of next month")+
  scale_x_discrete(labels = c("No default", "default")) +
  labs(title = "Default payment of next month",
       x = "Default payment",
       y = "Number of observations")
```

Since the number of records for 'default' is significantly lower than 'No default', the original dataset is somehow imbalance. 


```{r}
# address imbalance
# relevel
df_train1 <- mutate(df_train, default_pay=relevel(default_pay, 1))

df_test1 <- mutate(df_test, default_pay=relevel(default_pay, 1))
```

```{r}
# downsampling
set.seed(12)
fit1 <- train(default_pay ~ ., data=df_train1,
              method="glm", family=binomial(link="logit"),
              preProcess="medianImpute",
              trControl=trainControl(method="none",
                                     sampling="down"),
              na.action=na.pass)

confusionMatrix(predict(fit1, df_test1, na.action=na.pass),
                df_test1$default_pay,positive="1")

```


```{r}
# upsampling
set.seed(12)
fit1 <- train(default_pay ~ ., data=df_train1,
              method="glm", family=binomial(link="logit"),
              preProcess="medianImpute",
              trControl=trainControl(method="none",
                                     sampling="up"),
              na.action=na.pass)

confusionMatrix(predict(fit1, df_test1, na.action=na.pass),
                df_test1$default_pay,positive="1")
```





